Effects of the body force on the pedestrian and the evacuation dynamics

### 2020-03-04 paper.tex
algunas correcciones de ortografia

The Social Force Model (SFM) is a suitable model for describing crowd
behaviors under emotional stress. This research analyzes the role of the body
force in the original SFM. We focused on the parameter associated with the
body stiffness ($k_n$) and its impact on the pedestrian dynamics for two
different geometries: bottlenecks and corridors. Increasing $k_n$ produces
opposite effects on the crowd dynamics for each geometry: an increase of the
crowd velocity  for bottlenecks, and a decrease for corridors.  The former
reflects the fact that an increase in the stiffness reduces the overlap
between pedestrians and, as a consequence, the sliding friction is reduced.
This phenomenon reduces the number of blocking clusters close to the exit
door.  In the case of the corridor, instead, due to the confining walls, the
pedestrians get tight  into a lattice-like configuration due to space
limitations. The friction interaction with the walls determines the velocity
of the whole crowd along the corridor. Additionally, the corridor geometry
generates a flux slowing down for very crowded environments, as observed in
many real-life situations. We also explored the dimensionless parameters that
arose from the reduced-in-units equation of motion and tuned them to reproduce
the qualitative behavior of the empirical fundamental diagram.

@@

The Social Force Model (SFM) models the crowd dynamics  considering three kinds
of force: desired forces, social forces, and physical forces. In its original
version, the SFM, addresses two physical forces as essential:   the ``body
force'' and the ``sliding friction''. Both are  inspired by granular
interactions and were claimed to be necessary   for attaining the particular
effects in panicking crowds \cite{helbing_2000}.  The ``sliding friction''
actually proved to be an essential feature of the  ``faster-is-slower'' effect,
although the role of the ``body force'' appears,  at a first instance, not so
clear \cite{dorso_2005,dorso_2007,dorso_2011}. \\

The existence of a ``body force'' in the context of highly dense crowds (say, 5
to 10$\,$people/m$^2$) is a commonsense matter \cite{henein_2007,fruin_1993}.
Researchers, however, question the numerical setting for this force in  the SFM
context \cite{lakoba_2005}. As a matter of fact, the usual   set of parameters
provided by  Helbing prevents the excessive overlapping among pedestrians, but
it is known to  accomplish artificially high force levels
\cite{helbing_2000,lakoba_2005,langston_2006,lin_2017}. The force estimates from
the SFM appear to be remarkably higher with respect to the reported real life
data (say, an order of magnitude). The crowd motion simulations, however,
present realistic results \cite{lakoba_2005,langston_2006,dorso_2017}.  The
point seems to be that the SFM focuses on the ``collective behavior'' due  to
clogging, missing the ``individualistic'' perspective  of single pedestrians  or
very small groups \cite{helbing_2000,henein_2007,narain_2009}.  \\

@@

The fundamental diagram exhibits the flux behavior for either low density crowds
(with rare contacts between pedestrians) and highly dense crowds (dominated  by
two body interactions). In the second case, crowds experience a flux  slowing
down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}.
We may suspect that the  modeling of the  ``flux slowing down'' within the
context of the SFM will require the proper  setting of the (dimensionless)
controlling parameters. We have already examined this working hypothesis in
Ref.~\cite{dorso_2019}, but limiting the parameter exploration to  the sliding
friction, disregarding the body force.  We now widen the investigation on the
set of parameters to include the one  associated to the body force. We will
explore   the complex interplay between  the body force and the sliding friction
among pedestrians. Recall that the  interplay dynamics is not directly
controlled by the set of parameters, but  through dimensionless ``numbers'',
where the model parameters appear mixed  between each other. Thus, this step up
offers a challenge to the  ``individualistic'' meaning of the parameter's set.
\\

The paper is organized as follows. We first recall the available  experimental
values on the body force and the sliding friction (see Section
\ref{experimental}). Secondly, we introduce the reduced-in-units SFM and the
three dimensionless numbers that control the crowd dynamics (see Section
\ref{background}). We present our numerical simulations in Section
\ref{results}. For the sake of clarity, this Section is  separated into two
major parts: the bottleneck scenario and the corridor  scenario in Sections
\ref{bottleneck} and \ref{corridor}, respectively.  Section \ref{conclusions}
opens a detailed discussion from results in Section \ref{results} and resumes
our main conclusions.

@@

The reported data from Ref.~\cite{song_2019} correspond to the crowded
environment of the Beijing subway. This environment was not suitable for
providing information on the step distance $B$, but estimates for the  desired
speed and the body compressibility could be achieved. The reported  magnitude
$k_n$ assesses either the clothes and the body compressibility. The  final
value, though, is linearly related to the compression $x$. \\

@@

The maximum (realistic) compressions may be computed from the Hooke's relation
$F(x)=k_n(x)\,x$ and the compressibility $k_n(x)$ reported in
Table~\ref{table_data}.  An ``uncomfortable'' body force  $10\,$N $-$ $30\,$N
can address compressions in the  range of $0.030-0.055\,$m. Also, a ``hitting''
force of $60\,$N can address  compressions between $0.045$ and $0.065\,$m.
Besides, no reliable values for the sliding friction $k_t$ appears to be
available in the literature (to our knowledge).\\

@@

The Social Force Model (SFM) provides a necessary framework for simulating  the
collective dynamics of self-driven particles (\textit{i.e.} pedestrians).  The
pedestrians are considered to follow an equation of motion involving  either
``socio-psychological'' forces and physical forces (say, granular  forces). The
equation of motion for any pedestrian $i$ (of mass $m_i$) reads

@@

The pedestrians' own will is modeled by the desire force $\mathbf{f}_d$.  This
force stands for the acceleration (deceleration) required to move  at the
desired walking speed $v_d$. This involves, however, a  personal attitude that
makes him (her) appear more or less ``assertive''. As  mentioned in
Section~\ref{experimental}, the reaction time $\tau$ reflects for  this
attitude. Thus, the desire force is modeled as follows

@@

\subsection{Dimensionless numbers and comparison with empirical data}

>>>>

The Social Force Model (SFM) is a suitable model for describing crowd
behaviors under emotional stress. This research analyzes the role of the body
force in the original SFM. We focused on the parameter associated with the
body stiffness ($k_n$) and its impact on the pedestrian dynamics for two
different geometries: bottlenecks and corridors. Increasing $k_n$ produces
opposite effects on the crowd dynamics for each geometry: an increase of the
crowd velocity  for bottlenecks, and a decrease for corridors.  The former
reflects the fact that an increase in the stiffness reduces the overlap
between pedestrians and, as a consequence, the sliding friction is diminished.
This phenomenon reduces the number of blocking clusters close to the exit
door.  In the case of the corridor, instead, due to the confining walls, the
pedestrians get tight  into a lattice-like configuration due to space
limitations. The friction interaction with the walls determines the velocity
of the whole crowd along the corridor. Additionally, the corridor geometry
generates a flux slowing down for very crowded environments, as observed in
many real-life situations. We also explored the dimensionless parameters that
arose from the reduced-in-units equation of motion and tuned them to reproduce
the qualitative behavior of the empirical fundamental diagram.

@@

The Social Force Model (SFM) models the crowd dynamics  considering three kinds
of forces: desired forces, social forces, and physical forces. In its original
version, the SFM, addresses two physical forces as essential:   the ``body
force'' and the ``sliding friction''. Both are  inspired by granular
interactions and were claimed to be necessary   for attaining the particular
effects in panicking crowds \cite{helbing_2000}.  The ``sliding friction''
actually proved to be an essential feature of the  ``faster-is-slower'' effect,
although the role of the ``body force'' appears,  at a first instance, not so
clear \cite{dorso_2005,dorso_2007,dorso_2011}. \\

The existence of a ``body force'' in the context of highly dense crowds (say,
more than 5$\,$people/m$^2$) is a commonsense matter
\cite{henein_2007,fruin_1993}. Researchers, however, question the numerical
setting for this force in  the SFM context \cite{lakoba_2005}. As a matter of
fact, the usual   set of parameters provided by  Helbing prevents the excessive
overlapping among pedestrians, but it is known to  accomplish artificially high
force levels \cite{helbing_2000,lakoba_2005,langston_2006,lin_2017}. The force
estimates from the SFM appear to be remarkably higher with respect to the
reported real life data (say, an order of magnitude). The crowd motion
simulations, however, present realistic results
\cite{lakoba_2005,langston_2006,dorso_2017}.  The point seems to be that the SFM
focuses on the ``collective behavior'' due  to clogging, missing the
``individualistic'' perspective  of single pedestrians  or very small groups
\cite{helbing_2000,henein_2007,narain_2009}.  \\

@@

The fundamental diagram exhibits the flux behavior for either low density crowds
(with rare contacts between pedestrians) and highly dense crowds (dominated  by
two body interactions). In the second case, crowds experience a flux  slowing
down, but other behaviors are also possible \cite{helbing_2007,lohner_2018}. We
may suspect that the  modeling of the  ``flux slowing down'' within the context
of the SFM will require the proper  setting of the (dimensionless) controlling
parameters. This working hypothesis was already examined in
Ref.~\cite{dorso_2019}, but the parameter exploration was limited to  the
sliding friction, disregarding the body force. In this paper, we widen the
investigation on the set of parameters to include the one  associated to the
body force. We will explore   the complex interplay between  the body force and
the sliding friction among pedestrians. Recall that the  interplay dynamics is
not directly controlled by the set of parameters, but  through dimensionless
``numbers'', where the model parameters appear mixed  between each other. Thus,
this step up offers a challenge to the  ``individualistic'' meaning of the
parameter's set. \\

The paper is organized as follows. We first recall the available  experimental
values on the body force and the sliding friction (see Section
\ref{experimental}). Secondly, we introduce the reduced-in-units SFM and the
three dimensionless numbers that control the crowd dynamics (see Section
\ref{background}). We present our numerical simulations in Section
\ref{results}. For the sake of clarity, this Section is  separated into three
major parts: the bottleneck scenario, the corridor scenario and the comparison
with empirical data in Sections: \ref{bottleneck}, \ref{corridor} and
\ref{Dimensionless}, respectively.  Section \ref{conclusions} opens a detailed
discussion from results in Section \ref{results} and resumes our main
conclusions.

@@

The reported data from Ref.~\cite{song_2019} correspond to the crowded
environment of the Beijing subway. This environment was not suitable for
providing information on the step distance $B$, but estimates for the  desired
speed and the body compressibility could be achieved. The reported  magnitude
$k_n$ assesses either the clothes and the body compressibility. The  final
value, though, is linearly related to the compression (overlap) $x$. \\

@@

The maximum (realistic) overlap may be computed from the Hooke's relation
$F(x)=k_n(x)\,x$ and the compressibility $k_n(x)$ reported in
Table~\ref{table_data}.  An ``uncomfortable'' body force  $10\,$N $-$ $30\,$N
can address overlap values in the  range of $0.030-0.055\,$m. Also, a
``hitting'' force of $60\,$N can address  overlap values between $0.045$ and
$0.065\,$m. Besides, no reliable values for the sliding friction $k_t$ appears
to be available in the literature (to our knowledge).\\

@@

The Social Force Model (SFM) provides a necessary framework for simulating  the
collective dynamics of pedestrians (\textit{i.e.} self-driven particles). The
pedestrians are considered to follow an equation of motion involving  either
``socio-psychological'' forces and physical forces (say, granular  forces). The
equation of motion for any pedestrian $i$ (of mass $m_i$) reads

@@

The pedestrians' own will is modeled by the desire force $\mathbf{f}_d$.  This
force stands for the acceleration (deceleration) required to move  at the
desired walking speed $v_d$. This involves, however, a  personal attitude that
makes him (her) appear more or less ``assertive''. As  mentioned in
Section~\ref{experimental}, the reaction time $\tau$ reflects this
attitude. Thus, the desire force is modeled as follows

@@

\subsection{\label{Dimensionless}Dimensionless numbers and comparison
 with empirical data}
